---
title: "homework3"
author: xw2598
output: html_document
---
#problem2---write on the paper,this is a bonus
```{r p2-1}
library(tidyverse)
HeavySmoke = read.csv("HeavySmoke.csv",header = TRUE)
  HeavySmoke = HeavySmoke[, -1] 
#calculate the difference before and after 6 years quitting smoking
  HeavySmoke = mutate(HeavySmoke,diff = BMI_6yrs - BMI_base)
#show the table 
  HeavySmoke
#difference mean
  diff_mean = mean(HeavySmoke$diff)
#difference standard daviation
  diff_sd = sd(HeavySmoke$diff)
#Use one sample t-Test to test if the changes is equal to 0
  #Normality assumed
  #H0:mean difference = 0; H1: mean difference != 0
  #diff_mean = 3,36, diff_sd = 2.46, N = 10
  t_stat = diff_mean/(diff_sd/sqrt(10))
  #Compare the test statistics with the critical value, alpha=0.05
  qt(0.975,9)
  #t_stat>qt(9,0.975),reject H0
  p_value = 2*(1-pt(t_stat,9))
  tibble(
    d_bar = diff_mean,
    sd = diff_sd,
    t_stat = t_stat,
    critical_value = qt(0.975,9),
    alpha = 0.05,
    pvalue = p_value
  )
```
Interpretation:
Since t_stat = 4.30 >t9,0.975 and p-value < 10^-3, we reject the null hypothesis and conclude that there are sufficient evidence that women`s BMI has changed after 6 years quitting smoking.

```{r p2-2}
NeverSmoke = read.csv("NeverSmoke.csv",header = TRUE) 
NeverSmoke = NeverSmoke[,-1]
NeverSmoke = mutate(NeverSmoke, d2 = BMI_6yrs - BMI_base)
BmiChange = tibble(d1 = HeavySmoke$diff, d2 = NeverSmoke$d2)
#two sample
BmiChange
#using two sample t-test to test whether the difference of two samples` changes is equal to 0 
  #Normality assumed
  #H0: variances are equal; H1: variances are not equal  
d1_mean = mean(BmiChange$d1)
d1_sd = sd(BmiChange$d1)
d2_mean = mean(BmiChange$d2)
d2_sd = sd(BmiChange$d2)
#heavy smoke: n1 = 10, d1_bar = 3.36, sd1 = 2.46
#never smoke: n2 = 10, d2_bar = 1.55, sd2 = 2.28
F_stat = d1_sd^2/d2_sd^2
qf(0.975,9,9)
#F_stat = 1.55 < qf(0.975,9,9)
#There not enough evidences to declare a different variance, fail to reject H0
s = sqrt((9*d1_sd^2+9*d2_sd^2)/(10+10-2))
 #H0:d1_bar = d2_bar; H1:d1_bar != d2_bar
t2_stat = (d1_mean-d2_mean)/(s*sqrt(1/10+1/10))
qt(0.975,18)
#t2_stat<t18,0.975,fail to reject H0 and declare there is no enough evidencethe two sample has different mean.
tibble(
  d1_bar = d1_mean,
  sd1 =d1_sd,
  d2_bar = d2_mean,
  sd2 = d2_sd,
  estimate_s = s,
  F_stat = F_stat,
  F_Critical = qf(0.975,9,9),
  t_stat = t2_stat,
  t_Critical = qt(0.975,18)
)
```
Interpretation: On the significance level of 0.05, we fail to reject the H0 and conclude that there is no enough evidence that BMI changes are different between the two groups--women who quitted smoking and women who never smoked over the 6 years.

```{r p2_3}
# Construct a 95% CI for part2
CI = c((d1_mean - d2_mean) - qt(0.975,18)*s*sqrt(1/10+1/10), (d1_mean - d2_mean) + qt(0.975,18)*s*sqrt(1/10+1/10)) 
CI
```
Interpretation: We are 95% confident that the true mean difference of Bmi over 6 years changes between women who quitted smoking and who never smoked would be in (-0.421, 4.041)


Question a: 
I would design the study as a prespective cohort study since we have specific exposures and source population. 
In cohort study we needn`t to care much about randomizaton but we do need to care about subjects choosing. 
The location of choosing place may cause bias, for example, if we choose subjects from the hospital, the results may seem not true. Also, age, health condition, habits of exercises,nutrition differences can also cause bias. 
So we should make sure all of the factors that may cause BMI change among the two group subjects are the same.
Question b was written on the paper(attached in pdf)


#problem3
```{r Q1}
knee = read.csv("Knee.csv",header = TRUE)
descrip = tibble(
  var = c("Below", "Average", "Above" ),
  n = c(8,10,7),
  mean = c(mean(knee$Below,na.rm = TRUE),mean(knee$Average,na.rm = TRUE),mean(knee$Above,na.rm = TRUE)),
  sd =  c(sd(knee$Below,na.rm = TRUE), sd(knee$Average,na.rm = TRUE),sd(knee$Above,na.rm = TRUE))
  )

descrip
```
Comments:The mean decreasing order is :above,averange, below. This means people who have a worse physical status will need more time in physical therapy to get a successful rehabilitation.
The sd decreasing order is :averange, above, below. The smaller sd value is , the more convencing that  the group`s mean can represent its charater.

```{r Q2}
ins1 = na.omit(knee$Below)
ins2 = na.omit(knee$Average)
ins3 = na.omit(knee$Above)
time = c(ins1,ins2,ins3)
ind<-c(rep(1,length(ins1)),rep(2,length(ins2)),rep(3,length(ins3)))
new_data<-as.data.frame(cbind(time,ind))
# Perform an ANOVA test:
#H0:respect means of "average","below","above" groups are the same
#H1:respect means of "average","below","above" groups are not the same
new_data$ind = as.factor(new_data$ind)
res1<-aov(time~ind, data=new_data)
qf(0.99,2,22)
summary(res1)
```
Because F_stat = 19.28 > F0.99,2,22 = 5.72 , p-value = 1.45e-05, fail to reject H0.
At 0.01 significance level, we reject the null hypothesis and conclude that at least two of mean days required in physical therapy from three groups are different.
```{r, message=FALSE}
library(multcomp)
```

```{r Q3}

#pairwise comparisons 
#bonferroni
pairwise.t.test(new_data$time, new_data$ind,p.adj='bonferroni',conf.level = 0.01)
#Tukey
TukeyHSD(res1,conf.level = 0.01)
#Dunnett
test.dunnett=glht(res1, linfct=mcp(ind ="Dunnet"),conf.level = 0.01) 
summary(test.dunnett)
```
The bonferroni,Tukey and Dunnut results all indicate that there is a statistically significant difference between the "below" and "above"  and "averange" and "above". But Dunnet p-value is smaller than Turkey`s p-value than bonferroni.

Q4.There are extremely strong evidence that people who are above or average physical status will need obviously less time in physical therapy until successful rehabillitation than people who are below physical status. But whether there is difference between people who are above and average physical status, we are only 94% confident that above physical status people will require less time than average people.


#problem4
```{r}
library(datasets)
ncba = as.tibble(UCBAdmissions)
#overall proportion of men
p_male = sum(filter(ncba,Admit == "Admitted", Gender == "Male" )$n)/sum((filter(ncba,Gender == "Male" ))$n)
#overall number of men
n_man = sum(filter(ncba,Gender == "Male" )$n)
#overall proportion of women
p_female = sum(filter(ncba,Admit == "Admitted", Gender == "Female" )$n)/sum((filter(ncba,Gender == "Female" ))$n)
#overall number of women
n_women = sum(filter(ncba,Gender == "Female" )$n)
#approximation
n_women*p_female*(1-p_female)
n_man*p_male*(1-p_male)
#np(1-p)>5
#Normal Approximation of Binominal
var_men = p_male*(1-p_male)/n_man
var_women = p_female*(1-p_female)
#95%CI 
CI = c((p_male - p_female) - qnorm(0.975)*sqrt(var_men+var_women), (p_male - p_female) + qnorm(0.975)*sqrt(var_men+var_women)) 
CI
```
Interpretation: We are 95% confident that the true mean difference of difference proportions of  men vs women admitted at Berkeley would be in (-0.7597167,1.0430076)









