---
title: "homework3"
author: xw2598
output: html_document
---
#problem2---write on the paper,this is a bonus
```{r p2-1}
library(tidyverse)
HeavySmoke = read.csv("HeavySmoke.csv",header = TRUE)
  HeavySmoke = HeavySmoke[, -1] 
#calculate the difference before and after 6 years quitting smoking
  HeavySmoke = mutate(HeavySmoke,diff = BMI_6yrs - BMI_base)
#show the table 
  HeavySmoke
#difference mean
  diff_mean = mean(HeavySmoke$diff)
#difference standard daviation
  diff_sd = sd(HeavySmoke$diff)
#Use one sample t-Test to test if the changes is equal to 0
  #Normality assumed
  #H0:mean difference = 0; H1: mean difference != 0
  #diff_mean = 3,36, diff_sd = 2.46, N = 10
  t_stat = diff_mean/(diff_sd/sqrt(10))
  #Compare the test statistics with the critical value, alpha=0.05
  qt(0.975,9)
  #t_stat>qt(9,0.975),reject H0
  p_value = 2*(1-pt(t_stat,9))
  tibble(
    d_bar = diff_mean,
    sd = diff_sd,
    t_stat = t_stat,
    critical_value = qt(0.975,9),
    alpha = 0.05,
    pvalue = p_value
  )
```
Interpretation:
Since t_stat = 4.30 >t9,0.975 and p-value < 10^-3, we reject the null hypothesis and conclude that there are sufficient evidence that women`s BMI has changed after 6 years quitting smoking.

```{r p2-2}
NeverSmoke = read.csv("NeverSmoke.csv",header = TRUE) 
NeverSmoke = NeverSmoke[,-1]
NeverSmoke = mutate(NeverSmoke, d2 = BMI_6yrs - BMI_base)
BmiChange = tibble(d1 = HeavySmoke$diff, d2 = NeverSmoke$d2)
#two sample
BmiChange
#using two sample t-test to test whether the difference of two samples` changes is equal to 0 
  #Normality assumed
  #H0: variances are equal; H1: variances are not equal  
d1_mean = mean(BmiChange$d1)
d1_sd = sd(BmiChange$d1)
d2_mean = mean(BmiChange$d2)
d2_sd = sd(BmiChange$d2)
#heavy smoke: n1 = 10, d1_bar = 3.36, sd1 = 2.46
#never smoke: n2 = 10, d2_bar = 1.55, sd2 = 2.28
F_stat = d1_sd^2/d2_sd^2
qf(0.975,9,9)
#F_stat = 1.55 < qf(0.975,9,9)
#There not enough evidences to declare a different variance, fail to reject H0
s = sqrt((9*d1_sd^2+9*d2_sd^2)/(10+10-2))
 #H0:d1_bar = d2_bar; H1:d1_bar != d2_bar
t2_stat = (d1_mean-d2_mean)/(s*sqrt(1/10+1/10))
qt(0.975,18)
#t2_stat<t18,0.975,fail to reject H0 and declare there is no enough evidencethe two sample has different mean.
tibble(
  d1_bar = d1_mean,
  sd1 =d1_sd,
  d2_bar = d2_mean,
  sd2 = d2_sd,
  estimate_s = s,
  F_stat = F_stat,
  F_Critical = qf(0.975,9,9),
  t_stat = t2_stat,
  t_Critical = qt(0.975,18)
)
```
Interpretation: On the significance level of 0.05, we fail to reject the H0 and conclude that there is no enough evidence that BMI changes are different between the two groups--women who quitted smoking and women who never smoked over the 6 years.

```{r p2_3}
# Construct a 95% CI for part2
CI = c((d1_mean - d2_mean) - qt(0.975,18)*s*sqrt(1/10+1/10), (d1_mean - d2_mean) + qt(0.975,18)*s*sqrt(1/10+1/10)) 
CI
```
Interpretation: We are 95% confident that the true mean difference of Bmi over 6 years changes between women who quitted smoking and who never smoked would be in (-0.421, 4.041)

```{r p2_4}

#smokers: mu = 3.0, sd = 2.0
#never smokers: mu = 1.7, sd = 1.5
#2.5% significance level

# 5% significance level

```
Question a: 
I would design the study as a prespective cohort study since we have specific exposures and source population. 
In cohort study we needn`t to care much about randomizaton but we do need to care about subjects choosing. 
The location of choosing place may cause bias, for example, if we choose subjects from the hospital, the results may seem not true. Also, age, health condition, habits of exercises,nutrition differences can also cause bias. 
So we should make sure all of the factors that may cause BMI change among the two group subjects are the same.



#problem3
```{r}
knee = read.csv("Knee.csv",header = TRUE)
descrip = tibble(
  var = c("Below", "Average", "Above" ),
  n = c(8,10,7),
  mean = c(mean(knee$Below,na.rm = TRUE),mean(knee$Average,na.rm = TRUE),mean(knee$Above,na.rm = TRUE)),
  sd =  c(sd(knee$Below,na.rm = TRUE), sd(knee$Average,na.rm = TRUE),sd(knee$Above,na.rm = TRUE))
  )

descrip
```
Comments:The mean decreasing order is :above,averange, below. This means people who have a worse physical status will need more time in physical therapy to get a successful rehabilitation.
The sd decreasing order is :averange, above, below. The smaller sd value is , the more convencing that  the group`s mean can represent its charater.

```{r}
ins1 = na.omit(knee$Below)
ins2 = na.omit(knee$Average)
ins3 = na.omit(knee$Above)
insulin = c(ins1,ins2,ins3)
ind<-c(rep(1,length(ins1)),rep(2,length(ins2)),rep(3,length(ins3)))
new_data<-as.data.frame(cbind(insulin,ind))
# Perform an ANOVA test:
#H0:respect means of "average","below","above" groups are the same
#H1:respect means of "average","below","above" groups are not the same


```









